---
title: "Pokemon Data API Vignette"
author: "Collin Knezevich"
date: "6/19/2022"
output: github_document
---

# Pokemon Data API Vignette 

In this vignette, I will walk you through how to read data from [this Pokemon data API](https://pokeapi.co/). I will be creating functions to easily query data from the API, in such a way that it will be easy to call the function and query specific data.   

## Required Packages 

In order to query data from the Pokemon API, we will need the following packages: 

- `tidyverse` 
- `httr`
- `jsonlite` 

```{r, echo = FALSE, eval = TRUE}
# libraries 
library(tidyverse)
library(httr)
library(jsonlite)
```

## Functions 

### Pokemon Function 

The Pokemon API stores data about a specific Pokemon in two different places (both containing different info): a general "Pokemon" location, and a "Pokemon Species" location. This first function will query data from the former.   

The `returnPkmn` function takes either the ID number of a Pokemon, or the name of a Pokemon (as a string), as its input. It will return a data frame with the following relevant information from this location in the API: 
- *Name:* the name of the Pokemon 
- *ID:* the ID number of the Pokemon 
- *Type1:* the primary type of the Pokemon 
- *Type2:* the secondary type of the Pokemon, if applicable (will return NA if Pokemon does not have secondary type) 
- *HP:* base HP stat 
- *Attack:* base Attack stat 
- *Defense:* base Defense stat 
- *SpAttack:* base Special Attack stat 
- *SpDefense:* base Special Defense stat 
- *Speed:* base Speed stat

```{r}
returnPkmn <- function(idName){ 
  # set up URL for paste functions 
  startURL <- "https://pokeapi.co/api/v2/pokemon/" 
  endURL <- "/" 
  
  # if id specified 
  if (typeof(idName) == "double"){ 
    apiDat <- GET(paste0(startURL, as.character(idName), endURL)) 
    rawDat <- rawToChar(apiDat$content) 
    fullDat <- fromJSON(rawDat) 
  } 
  
  # if name specified 
  else if (typeof(idName) == "character"){ 
    name <- tolower(idName) 
    apiDat <- GET(paste0(startURL, name, endURL)) 
    rawDat <- rawToChar(apiDat$content) 
    fullDat <- fromJSON(rawDat) 
  } 
  
  # error control - if input improperly specified 
  else { 
    stop("Input incorrectly specified.") 
  } 
  
  # constructing data frame with relevant information 
  df <- data.frame(Name = fullDat$name, 
                   ID = fullDat$id, 
                   Type1 = fullDat$types$type$name[1],
                   Type2 = fullDat$types$type$name[2],
                   HP = fullDat$stats$base_stat[1], 
                   Attack = fullDat$stats$base_stat[2], 
                   Defense = fullDat$stats$base_stat[3], 
                   SpAttack = fullDat$stats$base_stat[4], 
                   SpDefense = fullDat$stats$base_stat[5], 
                   Speed = fullDat$stats$base_stat[6]) 
  return(df) 
}
```

### Species Function 

The `returnPkmnSpecies` function takes in either the ID number of a Pokemon, or the name of a Pokemon, as its input. It will return a data frame with the following particularly relevant information from the API:   
- *Name:* the name of the Pokemon 
- *ID:* the ID number of the Pokemon 
- *Gen:* the generation in which the Pokemon first appeared 
- *PreEvo:* the Pokemon from which the queried Pokemon evolved from (returns NULL if no pre-evolution) 
- *EggGrp:* the egg group of the Pokemon 
- *Legendary:* a boolean to say whether or not the Pokemon is a legendary 
- *Mythical:* a boolean to say whether or not the Pokemon is a mythical

```{r}
# Pokemon species function 
returnPkmnSpecies <- function(idName){ 
  # set up URL for paste functions 
  startURL <- "https://pokeapi.co/api/v2/pokemon-species/"
  endURL <- "/"
  
  # if id specified 
  if (typeof(idName) == "double"){ 
    apiDat <- GET(paste0(startURL, as.character(idName), endURL)) 
    rawDat <- rawToChar(apiDat$content) 
    fullDat <- fromJSON(rawDat) 
  } 
  
  # if name specified 
  else if (typeof(idName) == "character"){ 
    name <- tolower(idName) 
    apiDat <- GET(paste0(startURL, name, endURL)) 
    rawDat <- rawToChar(apiDat$content) 
    fullDat <- fromJSON(rawDat) 
  } 
  
  # error control - if input improperly specified 
  else { 
    stop("Input incorrectly speficied.")
  } 
  
  # constructing data frame with relevant information 
  df <- data.frame(Name = fullDat$name, 
                   ID = fullDat$id,  
                   Gen = fullDat$generation$name, 
                   PreEvo = fullDat$evolves_from_species$name, 
                   EggGrp = fullDat$egg_groups$name, 
                   Legendary = fullDat$is_legendary, 
                   Mythical = fullDat$is_mythical) 
  return(df)
}
```
